---
title: "Final Project"
author: "Jie Sun"
date: "4/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## import the data
```{r}
#install.packages(SASxport) #Install SASxport package for reading xpt file
#install.packages('VIM') #Install VIM package for plotting missing data patterns
library(SASxport)
library(VIM) 

# read data
BRFSS2018<-read.xport("/Users/jiesun/Downloads/LLCP2018.xpt ")

# drop unnecessary columns
BRFSS <- BRFSS2018[c("ADSLEEP","EMPLOY1","EXERANY2","MENTHLTH","MARITAL","DRNK3GE5")]
```
```{r}
# Look at missing data patterns using the md.pattern function
md.pattern(BRFSS, rotate.names=TRUE)
```

## data management
```{r}
# look at variables
table(BRFSS$ADSLEEP) # number of days have sleep disorder over 2 weeks
table(BRFSS$EMPLOY1) # employment status
table(BRFSS$EXERANY2) # exercise in past 30 days (yes or no)
table(BRFSS$MENTHLTH) # number of days have mental health problems in the past 30 days
table(BRFSS$MARITAL) # marital status
table(BRFSS$DRNK3GE5) # alcohol consumption

# recode "don't know", "refused", "blank" to missing values
BRFSS$ADSLEEP[BRFSS$ADSLEEP > 14 & BRFSS$ADSLEEP !=88] <- NA
BRFSS$EMPLOY1[BRFSS$EMPLOY1 > 8] <- NA
BRFSS$EXERANY2[BRFSS$EXERANY2 > 2] <- NA
BRFSS$MENTHLTH[BRFSS$MENTHLTH > 30 & BRFSS$MENTHLTH !=88] <- NA
BRFSS$MARITAL[BRFSS$MARITAL > 6] <- NA
BRFSS$DRNK3GE5[BRFSS$DRNK3GE5 > 76] <- NA

# Look at missing data patterns using the md.pattern function
md.pattern(BRFSS, rotate.names=TRUE)

# remove missing data
BRFSS <- BRFSS[!is.na(BRFSS$ADSLEEP), ]
BRFSS <- BRFSS[!is.na(BRFSS$EMPLOY1), ]
BRFSS <- BRFSS[!is.na(BRFSS$EXERANY2), ]
BRFSS <- BRFSS[!is.na(BRFSS$MENTHLTH), ]
BRFSS <- BRFSS[!is.na(BRFSS$MARITAL), ]
BRFSS <- BRFSS[!is.na(BRFSS$DRNK3GE5), ]

# recode none(88) to 0 for ADSLEEP and MENTHLTH
BRFSS$ADSLEEP[BRFSS$ADSLEEP == 88] <- 0
BRFSS$MENTHLTH[BRFSS$MENTHLTH == 88] <- 0

# recode categorical variables
BRFSS$EXERANY2<-factor(BRFSS$EXERANY2, levels=c(1:2), labels=c("yes", "no"))
BRFSS$EMPLOY1_re[BRFSS$EMPLOY1<3] <- 0
BRFSS$EMPLOY1_re[(BRFSS$EMPLOY1>4 & BRFSS$EMPLOY1<7)] <- 1
BRFSS$EMPLOY1_re[(BRFSS$EMPLOY1 == 3 | BRFSS$EMPLOY1 == 4 | BRFSS$EMPLOY1 == 7 | BRFSS$EMPLOY1 == 8)] <- 2
BRFSS$EMPLOY1_re<-factor(BRFSS$EMPLOY1_re, levels=c(0:2), labels=c("employed", "out of work", "student or homemaker"))
BRFSS$MARITAL_re[BRFSS$MARITAL == 1] <- 0
BRFSS$MARITAL_re[BRFSS$MARITAL > 1] <- 1
BRFSS$MARITAL_re<-factor(BRFSS$MARITAL_re, levels=c(0:1), labels=c("married", "single"))
# Check data
View(BRFSS)

# Check shape of distribution
d <- density(BRFSS$ADSLEEP)
plot(d, xlim=c(0,14))
hist(BRFSS$ADSLEEP, breaks=c(0,2,4,6,8,10,12,14))
```
# descriptive analysis
```{r}
table(BRFSS$EMPLOY1_re)
table(BRFSS$EXERANY2)
table(BRFSS$MARITAL_re)
mean(BRFSS$ADSLEEP)
sd(BRFSS$ADSLEEP)
mean(BRFSS$MENTHLTH)
sd(BRFSS$MENTHLTH)
mean(BRFSS$DRNK3GE5)
sd(BRFSS$DRNK3GE5)
```

# run first poisson regression
```{r}
# the first model contains physical activity, mental health, and employment status, which are 3 predictors that we're interested
m1 <- glm(ADSLEEP ~ EXERANY2 + MENTHLTH + EMPLOY1_re, family ="poisson", data = BRFSS)
summary(m1)
```

# Cameron and Trivedi (2009) recommended using robust standard errors for the parameter estimates to control for mild violation of the distribution assumption that the variance equals the mean. We use R package sandwich below to obtain the robust standard errors and calculated the p-values accordingly. Together with the p-values, we have also calculated the 95% confidence interval using the parameter estimates and their robust standard errors.
```{r}
#install.packages(msm)
#install.packages(sandwich)
library(msm)
library(sandwich)
cov.m1 <- vcovHC(m1, type="HC0")
std.err <- sqrt(diag(cov.m1))
r.est <- cbind(Estimate= coef(m1), "Robust SE" = std.err,
"Pr(>|z|)" = 2 * pnorm(abs(coef(m1)/std.err), lower.tail=FALSE),
LL = coef(m1) - 1.96 * std.err,
UL = coef(m1) + 1.96 * std.err)
```

# We want to present the regression results as incident rate ratios and their standard errors, together with the confidence interval. To compute the standard error for the incident rate ratios, we use the Delta method. To this end, we make use the function deltamethod implemented in R package msm.
```{r}
s <- deltamethod(list(~ exp(x1), ~ exp(x2), ~ exp(x3), ~ exp(x4), ~ exp(x5)), coef(m1), cov.m1)
# exponentiate old estimates dropping the p values
rexp.est <- exp(r.est[, -3])
# replace SEs with estimates for exponentiated coefficients
rexp.est[, "Robust SE"] <- s

rexp.est
```

# run second poisson regresson
```{r}
# marital status and alcohol usage are included as confounders in this model to be adjusted
m2 <- glm(ADSLEEP ~ EXERANY2 + MENTHLTH + EMPLOY1_re + MARITAL_re + DRNK3GE5, family ="poisson", data = BRFSS)
summary(m2)

# calcuate IRR
cov.m2 <- vcovHC(m2, type="HC0")
std.err.2 <- sqrt(diag(cov.m2))
r.est.2 <- cbind(Estimate= coef(m2), "Robust SE" = std.err.2,
"Pr(>|z|)" = 2 * pnorm(abs(coef(m2)/std.err.2), lower.tail=FALSE),
LL = coef(m2) - 1.96 * std.err.2,
UL = coef(m2) + 1.96 * std.err.2)

s2 <- deltamethod(list(~ exp(x1), ~ exp(x2), ~ exp(x3), ~ exp(x4), ~ exp(x5), ~ exp(x6), ~ exp(x7)), coef(m2), cov.m2)

## exponentiate old estimates dropping the p values
rexp.est.2 <- exp(r.est.2[, -3])
## replace SEs with estimates for exponentiated coefficients
rexp.est.2[, "Robust SE"] <- s2

rexp.est.2
```

# What about overdispersion: Use quasi-poisson to get dispersion parameter
```{r, eval=FALSE}
#get overdispersion parameter from model results 
pr <- residuals(m2,"pearson")
phi <- sum(pr^2)/df.residual(m2)
round(c(phi,sqrt(phi)),4) #sqrt of phi is the number that the SEs will be multiplied by to adjust them for overdispersion
####extra####

## quasi-Poisson to allow the scale parameter to change from 1. Show the dispersion parameter. If the dispersion parameter in quasi-Poisson is >1 we have overdispersion
model.2q <- glm(ADSLEEP ~ EXERANY2 + MENTHLTH + EMPLOY1_re + MARITAL_re + DRNK3GE5, family = quasipoisson, data = BRFSS)
summary(model.2q)
```
The dispersion parameter in quasi-Poisson is >1, so we have overdispersion.
We use robust standard errors above, which can correct for SEs overdispersion.

# We can also use negbin to overcome overdispersion 
```{r, eval=FALSE}
# Load MASS
library(MASS)
# Negative binomial regression
m2.nb <- glm.nb(ADSLEEP ~ EXERANY2 + MENTHLTH + EMPLOY1_re + MARITAL_re + DRNK3GE5, data = BRFSS)
summary(m2.nb)
```

# We can also decide which model is better to use (Poisson vs. negbin by comparing models using the liklihood ratio test)
```{r, eval=FALSE}
library(lmtest) #model comparison
lrtest(m2, m2.nb)
```
Since p-value<0.001, negbin model is better to use.

